\textmd{\hmwkClass:\ \hmwkTitle}\\ \small{Due\ on\ \hmwkDueDate}

### 2017-05-08 paper/main.tex
fixed merge conflict

\title{
    \textmd{\hmwkClass:\ \hmwkTitle}\\
    % \normalsize{\hmwkClassInstructor}\\
    % \small{Issued\ on\ \hmwkIssueDate}\\
    \small{Due\ on\ \hmwkDueDate}
}
\author{\hmwkAuthorName}
\date{}
\begin{document}
\maketitle
\section{Introduction}
motivation: features of physics engine that incorporate ideas from class (e.g. generic operator, hierarchy, propagator, language for language (high--low)), cool visualization of concepts from class

@@

The coercion method is 
\begin{verbatim}
(define (coerce a size-reference)
  (if (vector? a)
    a
    (make-vector (vector-length size-reference) a)))
\end{verbatim}
The rest of the infrastructure for generic procedures was adapted from pset 4, but pset 4's generic arithmetic is not compatible with that of pset 3. To fix this issue, we first loaded the relevant pset 3 dependencies, installed a generic arithmetic with vector operations, redefined arithmetic operations to be the ones from that generic arithmetic (such as with \texttt{(define * (access * arith-environment))}), and then loaded the pset 4 generic procedure infrastructure on top.

\section{Physics Engine}
\subsection{Object Types}
There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}. \texttt{thing} refers to all physical objects in the system. There is a hierarchy of physical objects that have additional properties. For example, type \texttt{point-charge} has an additional property \texttt{charge} and experiences electric forces while type \texttt{ball} experiences only gravity. Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself. \texttt{interaction} refers to forces exerted on \texttt{thing}s. \texttt{interaction} includes a procedure that calculates the force as well as \texttt{influences}, a list of \texttt{thing}s that contribute to that force. Similar to \texttt{thing}, subtypes of \texttt{interaction}s form a hierarchy. Lastly, \texttt{world} keeps a list of all \texttt{thing}s and an \texttt{update} call to the \texttt{world} calls on each \texttt{thing} to update itself.

@@

\subsection{Output Display}
Our engine updates each object every timestep and re-renders them on the screen:
\begin{verbatim}
(define (run-engine world steps)
  (reset-graphics)
  (if (> steps 0)
    (begin 
      (for-each 
        (lambda (thing)
            (newline)
            (display (cons (get-name thing) (get-position thing)))  ; for debugging
            (render thing))
        (get-world-all-things world))
      (update-world world)
    (run-engine world (- steps 1)))))
\end{verbatim}

@@

\subsection{Magentism}
% \begin{samepage}
Let's add electric forces between point charges into the physics engine.
{\small\begin{verbatim}
(define (make-electric-force point-charge all-point-charges)
  (define (procedure point-charge influences)
    (sum (map (lambda (influence)
                (let* ((q1 (get-point-charge-charge point-charge))
            		       (q2 (get-point-charge-charge influence))
            		       (epsilon0 8.854e-12) ; vacuum permittivity
                       ;; vector between influence and point charge
            		       (v (- (get-position influence)
                             (get-position point-charge)))
                       ;; distance between influence and point charge
            		       (r (magnitude v))
                       ;; unit vector * -1 to account for opposite signs
                       ;; attract & same signs repel
            		       (u (* -1 (/ v r)))
            		       (melec (/ (* q1 q2)  ; magnitude of electric force
                                 (* 4 pi epsilon0 (square r)))))
                  (* u melec))) ; multiply unit vector by magnitude of force
              influences)))
  (let ((influences (delq point-charge all-point-charges)))
    (make-interaction electric-force? 'electric-force procedure influences)))
\end{verbatim}}
% \end{samepage}

% \begin{samepage}
This is a simple world with two point charges repelling each other.
{\small\begin{verbatim}
(define (charges-1)
  (define w (make-world "world"))
  (define m1 (make-point-charge "charge1" 1e-3 1 #(-30 -30) #(0 0) "red"))
  (define m2 (make-point-charge "charge2" 1e-3 1 #(30 30) #(0 0) "red"))

@@

% \begin{samepage}
This world contains two positively charged masses (blue) and two negatively charged masses (red).
{\small\begin{verbatim}
(define (charges-2)
  (define w (make-world "world"))
  (define m1 (make-point-charge "charge1" 1e-3 1 #(-100 -100) #(0 0) "red"))
  (define m2 (make-point-charge "charge2" -1e-3 1 #(100 100) #(0 0) "blue"))
  (define m3 (make-point-charge "charge3" 1e-3 1 #(-100 100) #(0 0) "red"))
  (define m4 (make-point-charge "charge4" -1e-3 1 #(100 -100) #(0 0) "blue"))

@@

% \begin{samepage}
We can also combine gravity and electric charge. Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and electric force. As figure \ref{figure:charges} (right) shows, this charged solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.
{\small\begin{verbatim}
(define (charged-solar-system)
  (define w (make-world "world"))
  (define s (make-point-charge "sun" 3e-3 1e15 #(0 0) #(0 0) "blue"))
  (define b1 (make-point-charge "ball1" -1e-1 1e5 #(100 100) #(-15.361 15.361) "red"))
  (define b2 (make-point-charge "ball2" -1e-1 1e5 #(110 110) #(-15.361 15.361) "red"))
  (define b3 (make-point-charge "ball3" -1e-1 1e5 #(120 120) #(-15.361 15.361) "red"))
  (define b4 (make-point-charge "ball4" -1e-1 1e5 #(130 130) #(-15.361 15.361) "red"))
  (define b5 (make-point-charge "ball5" -1e-1 1e5 #(140 140) #(-15.361 15.361) "red"))
  (define b6 (make-point-charge "ball6" -1e-1 1e5 #(150 150) #(-15.361 15.361) "red"))

@@

\section{Extensibility of the System}
Possible future extensions and high level explanation of how they would be implemented (through minimal change to current system)

>>>>

\title{
    \textmd{\hmwkClass:\ \hmwkTitle}\\
    % \normalsize{\hmwkClassInstructor}\\
    % \small{Issued\ on\ \hmwkIssueDate}\\
    \small{Due\ on\ \hmwkDueDate}
}
\author{\hmwkAuthorName}
\date{}
\begin{document}
\maketitle
\section{Introduction}

motivation: features of physics engine that incorporate ideas from class (e.g.
generic operator, hierarchy, propagator, language for language (high--low)),
cool visualization of concepts from class

@@

The coercion method is 
\begin{verbatim}
(define (coerce a size-reference)
  (if (vector? a)
    a
    (make-vector (vector-length size-reference) a)))
\end{verbatim}

The rest of the infrastructure for generic procedures was adapted from pset 4,
but pset 4's generic arithmetic is not compatible with that of pset 3. To fix
this issue, we first loaded the relevant pset 3 dependencies, installed a
generic arithmetic with vector operations, redefined arithmetic operations to
be the ones from that generic arithmetic (such as with \texttt{(define *
(access * arith-environment))}), and then loaded the pset 4 generic procedure
infrastructure on top.

\section{Physics Engine}
\subsection{Object Types}

There are three types of objects in the system: \texttt{thing}, \texttt{interaction}, and \texttt{world}. \texttt{thing} refers to all physical objects in the system. There is a hierarchy of physical objects that have additional properties. For example, type \texttt{point-charge} has an additional property \texttt{charge} and experiences electric forces while type \texttt{ball} experiences only gravity. Each \texttt{thing} keeps a list of \texttt{interaction} that is used to update itself. \texttt{interaction} refers to forces exerted on \texttt{thing}s. \texttt{interaction} includes a procedure that calculates the force as well as \texttt{influences}, a list of \texttt{thing}s that contribute to that force. Similar to \texttt{thing}, subtypes of \texttt{interaction}s form a hierarchy. Lastly, \texttt{world} keeps a list of all \texttt{thing}s and an \texttt{update} call to the \texttt{world} calls on each \texttt{thing} to update itself.

@@

\subsection{Output Display}

Our engine updates each object every timestep and re-renders them on the
screen:

\begin{verbatim}
(define (run-engine world steps)
  (reset-graphics)
  (if (> steps 0)
    (begin 
      (for-each 
        (lambda (thing)
            (newline)
            (display (cons (get-name thing) (get-position thing)))  ; for debugging
            (render thing))
        (get-world-all-things world))
      (update-world world)
    (run-engine world (- steps 1)))))
\end{verbatim}

@@

\subsection{Electromagentism}
% \begin{samepage}

Let's add electric forces between point charges into the physics engine.

{\small\begin{verbatim}
(define (make-electric-force point-charge all-point-charges)
  (define (procedure point-charge influences)
    (sum (map (lambda (influence)
                (let* ((q1 (get-point-charge-charge point-charge))
            		       (q2 (get-point-charge-charge influence))
            		       (epsilon0 8.854e-12) ; vacuum permittivity
                       ;; vector between influence and point charge
            		       (v (- (get-position influence)
                             (get-position point-charge)))
                       ;; distance between influence and point charge
            		       (r (magnitude v))
                       ;; unit vector * -1 to account for opposite signs
                       ;; attract & same signs repel
            		       (u (* -1 (/ v r)))
            		       (melec (/ (* q1 q2)  ; magnitude of electric force
                                 (* 4 pi epsilon0 (square r)))))
                  (* u melec))) ; multiply unit vector by magnitude of force
              influences)))
  (let ((influences (delq point-charge all-point-charges)))
    (make-interaction electric-force? 'electric-force procedure influences)))
\end{verbatim}}
% \end{samepage}

% \begin{samepage}

This is a simple world with two point charges repelling each other.

{\small\begin{verbatim}
(define (charges-1)
  (define w (make-world "world"))
  (define m1 (make-point-charge "charge1" 1e-3 1 #(-30 -30) #(0 0) "red"))
  (define m2 (make-point-charge "charge2" 1e-3 1 #(30 30) #(0 0) "red"))

@@

% \begin{samepage}

This world contains two positively charged masses (blue) and two negatively
charged masses (red).

{\small\begin{verbatim}
(define (charges-2)
  (define w (make-world "world"))
  (define m1 (make-point-charge "charge1" 1e-3 1 #(-100 -100) #(0 0) "red"))
  (define m2 (make-point-charge "charge2" -1e-3 1 #(100 100) #(0 0) "blue"))
  (define m3 (make-point-charge "charge3" 1e-3 1 #(-100 100) #(0 0) "red"))
  (define m4 (make-point-charge "charge4" -1e-3 1 #(100 -100) #(0 0) "blue"))

@@

% \begin{samepage}

We can also combine gravity and electric charge. Here, we created a solar-system-like world where each of the planets repel each other and the sun attracts the planets via gravity and electric force. As figure \ref{figure:charges} (right) shows, this charged solar system exhibits wider and diverging orbits compared to figure \ref{figure:gravity} (right), the solar system with only gravity.

{\small\begin{verbatim}
(define (charged-solar-system)
  (define w (make-world "world"))
  (define s (make-point-charge "sun" 3e-3 1e15 #(0 0) #(0 0) "blue"))
  (define b1 (make-point-charge "ball1" -1e-1 1e5 #(100 100) #(-15.361 15.361) "red"))
  (define b2 (make-point-charge "ball2" -1e-1 1e5 #(110 110) #(-15.361 15.361) "red"))
  (define b3 (make-point-charge "ball3" -1e-1 1e5 #(120 120) #(-15.361 15.361) "red"))
  (define b4 (make-point-charge "ball4" -1e-1 1e5 #(130 130) #(-15.361 15.361) "red"))
  (define b5 (make-point-charge "ball5" -1e-1 1e5 #(140 140) #(-15.361 15.361) "red"))
  (define b6 (make-point-charge "ball6" -1e-1 1e5 #(150 150) #(-15.361 15.361) "red"))

@@

\section{Extensibility of the System}

Possible future extensions and high level explanation of how they would be
implemented (through minimal change to current system)
